---
title: "Lab 7s"
author: "Olivia Kelly"
format:
  html:
    toc: true
    toc_float: true
    code-fold: true
    embed-resources: true
execute:
  warning: false 
  message: false 
---

### Background 

```{r}
library(neonUtilities)
library(tidyverse)
library(lubridate)
library(DT)
library(viridis)
```

```{r}
library(tidyverse)
library(lubridate)
library(DT)
library(viridis)
library(janitor)
library(plotly)
library(respirometry)
```

```{r}
soilChem <- loadByProduct(
  dpID='DP1.10086.001',
  startdate = "2023-01",
  enddate = "2023-12",
  check.size = FALSE,
  package='expanded')
```

```{r}
soilChem$sls_metagenomicsPooling[5,'genomicsSampleID']
```

```{r}
soilChem$sls_metagenomicsPooling[5,'collectDate']
```

```{r}
soilChem$sls_metagenomicsPooling$genomicsPooledIDList[5]
```

```{r}
# View(soilChem$sls_soilChemistry # No soil chemistry data for HARV
soilChem$sls_soilMoisture |> 
  filter(sampleID == "HARV_013-O-39-20230704") |> 
  select(sampleID, soilMoisture)
```

```{r}
# split up the pooled list into new columns
genomicSamples <- soilChem$sls_metagenomicsPooling |>
  tidyr::separate(genomicsPooledIDList, into=c("first","second","third"),sep="\\|",fill="right") |>
  dplyr::select(genomicsSampleID,first,second,third)

head(genomicSamples)
```

```{r}
genSampleExample <- genomicSamples |> 
  tidyr::pivot_longer(cols=c("first","second","third"),values_to = "sampleID") |>
  dplyr::select(sampleID,genomicsSampleID) |>
  drop_na()
```

```{r}
chemEx <- soilChem$sls_soilMoisture |>
  dplyr::select(sampleID,soilMoisture)

## now combine the tables 
combinedTab <- left_join(genSampleExample,chemEx, by = "sampleID") |> drop_na()

## Show the data table
head(combinedTab)
```

```{r}
soilpH_Example <- soilChem$sls_soilpH %>%
  dplyr::filter(sampleID %in% combinedTab$sampleID) %>%
  dplyr::select(sampleID,soilInWaterpH,soilInCaClpH)

# now join with the existing table
combinedTab_pH <- left_join(combinedTab,soilpH_Example, by = "sampleID")

# and the final
head(combinedTab_pH)
```

```{r}
# Remember to install the respirometry package
library(respirometry)

genome_groups_pH <- combinedTab_pH %>%
  group_by(genomicsSampleID) %>%
  summarize_at(c("soilInWaterpH","soilInCaClpH"), mean_pH) 

head(genome_groups_pH)
```

```{r}
genome_groups_all_mean <- combinedTab_pH %>%
  group_by(genomicsSampleID) %>%
  {left_join(
    summarize_at(.,vars("soilMoisture"), mean),
    summarize_at(.,vars("soilInWaterpH","soilInCaClpH"), mean_pH)
  )}

head(genome_groups_all_mean)
```

## 2. Environmental Correlations

-   Soil chemistry vs. taxonomy: Correlate pH, moisture, temperature, and elevation with microbial diversity or specific taxa.

-   Gradient analysis: Identify taxa enriched along environmental gradients (e.g., moisture or pH).

-   Canonical Correspondence Analysis (CCA): Link community composition to environmental variables.

Using copoilet I was given the following, I have to keep wokring through and asking more questions because there was some errors in the codes thats why they're not in code chunks becasuse I needed it to render.

```{r}
class(df)

```

df \<- read.csv("your_data.csv") \# or however you imported it

library(dplyr) library(vegan)

# Create the summarized dataframe

diversity_df \<- df %\>% group_by(genomicsSampleID) %\>% summarize( shannon = diversity(abundance, index = "shannon"), \# replace 'abundance' with your counts column soil_pH = mean(soilInWaterpH, na.rm = TRUE), soil_moisture = mean(soilMoisture, na.rm = TRUE), soil_temp = mean(soilTemperature, na.rm = TRUE), elevation = mean(elevation, na.rm = TRUE) )

# Now you can use diversity_df

cor.test(diversity_df$soil_pH, diversity_df$shannon) cor.test(diversity_df$soil_moisture, diversity_df$shannon)

library(ggplot2)

ggplot(diversity_df, aes(x = soil_pH, y = shannon)) + geom_point() + geom_smooth(method = "lm") + theme_minimal()

# Assuming diversity_df has shannon + soil variables

cor.test(diversity_df$soil_pH, diversity_df$shannon) cor.test(diversity_df$soil_moisture, diversity_df$shannon) cor.test(diversity_df$soil_temp, diversity_df$shannon) cor.test(diversity_df$elevation, diversity_df$shannon)
